\documentclass{amsart}

\begin{document}
\begin{center}
 {\Large\bf PMA101 --- Response to checker's comments}
\end{center}

\begin{itemize}
 \item Students have been informed several times about the format, and
  will be informed again.
 \item I have put the word ``all'' in bold.
 \item The information already given to students mentions ``Section
  A'' and ``Section B'', so I will retain this terminology.
 \item I have 
  \begin{itemize}
   \item moved A13 into section B;
   \item added a new question (A10) worth 2 marks;
   \item added an extra mark to each of A6, A14 and A15;
   \item simplified B1 and taken away two marks;
   \item replaced the old B3 by a different question, now called B4.
  \end{itemize}
 \item A3: I have told the students to simplify their answer. 
 \item A4: I have used $\log$ instead of $\ln$ consistently in the
  notes and lectures.
 \item A12,A13 (was A11,A12): I will write a mark scheme for
  integration by parts if necessary.  The method in the solutions has
  been used in tutorials and lectures and I expect that the majority
  of students will follow it.
 \item B3 (was A13): I will tell the students that they should know
  $\int\log(x)\,dx$.  Nonetheless, I have allocated an extra mark.
 \item A14,A15: I have allocated an extra mark for each of these.
 \item B1: I have simplified the question as suggested, and moved two
  marks elsewhere.
 \item B2: I have subtracted one mark.
 \item old B3: I have removed this.
\end{itemize}

\bibliography{bibdata,strickland}
\end{document}
